Capabilities, MMIO and Robust Safety

### 2020-04-19 capability-notes/driver-notes/main.tex
Driver: small note on non-atomic invariants

@@

\subsubsection{Concrete boot and driver code}
\label{subsec:driverimpl}
Now we have a closer look at the concrete example we would like to verify,
including what its boot code looks like. Our goal is to prove that this concrete
example satisfies \X{bootSpec} (alternatively, \textsc{BootCodeOk}) and this
last robust safety theorem.

@@

We have to tie this integer $n$ to the number of I/O events
present in the trace $t$, otherwise it will prove impossible to verify our
driver. Concretely, the invariant we will verify our driver under changes from
$\trInv[|t| \le 1000]$ to
$\trInvNp[\pure{|t| \le 1000} \ast
  \exists n\ldotp l_n \mapsto n \ast n = |t| ]$, where we defined
%
\[
  \begin{array}{lcl}
    \trInvNp[P_{\X{np}}] & \eqdef & \knowInv{\iota}{\exists t.\; \trace(t) \ast P_{\X{np}}(t)}
   \end{array}
\]

Actually, there are two options, where the second one is probably slightly more modular:
\[
  \begin{array}{lcl}

@@

(i.e., the predicate over the trace does not need to be pure any longer).
Concretely, this means that
instead of having a pure predicate $P$ in our robust safety theorem, we should
also allow non-pure predicates $P_{\X{np}}$ that entail $P$, i.e. $\forall t
\ldotp P_{\X{np}}(t) \wand \pure{P(t)} $, since we might have to enforce
other invariants relating to local state over the trace as well.

>>>>

Lastly, note how the trace invariant that appears in the boot specification is not an atomic invariant, even though atomic invariants might be involved in the verification of the driver if it has state (see e.g.\ the counter example below). We are obliged to have a regular invariant for the predicate over the trace, since otherwise adequacy breaks; even for infinite executions, we have to be able to show that any output values satisfy $P$ at \emph{any} point during execution.

@@

\subsubsection{Concrete boot and driver code}
\label{subsec:driverimpl}
Now we have a closer look at the concrete example we would like to verify,
including what its boot code looks like. Our goal is to prove that this concrete
example satisfies \X{bootSpec} and hence the above robust safety theorem.

@@

We have to tie this integer $n$ to the number of I/O events
present in the trace $t$, otherwise it will prove impossible to verify our
driver. Concretely, the invariant we will verify our driver under changes from

\newcommandx{\NaInvT}[1][1 = \top]{[\X{NaInv}: \X{driver\_nais}.#1]}
\newcommandx{\NaInvI}[2][1 = \iota]{\operatorname{NaInv}^{#1}(#2)}
$\trInv[|t| \le 1000]$ to
$\trInvNp[\exists n \ldotp  \pure{|t| = n \mathrel{\land} n \le 1000} \mathrel{\ast} \ownGhost{\gamma_n}{\authfull n}] \mathrel{\ast} \NaInvI[\iota]{\exists n\ldotp l_n \mapsto n \ast \ownGhost{\gamma_n}{\authfrag n}]}  $, where we defined
%
\[
  \begin{array}{lcl}
    \trInvNp[P_{\X{np}}] & \eqdef & \knowInv{\iota}{\exists t.\; \trace(t) \ast P_{\X{np}}(t)}
   \end{array}
\]
(i.e., the predicate over the trace does not need to be pure any longer).
Concretely, this means that
instead of having a pure predicate $P$ in our robust safety theorem, we should
also allow non-pure predicates $P_{\X{np}}$ that entail $P$, i.e. $\forall t
\ldotp P_{\X{np}}(t) \wand \pure{P(t)} $, since we might have to enforce
other invariants relating to local state over the trace as well.

Note that the following two formalizations do not work, since the trace and counter value are not updated at the same time in the code:
\[
  \begin{array}{lcl}

@@

Using  a non-atomic invariant in the first of these two formalizations would not work either, as it would break the adequacy (as stated earlier); we would not know that the safety property inside $P_{\X{np}}$ was upheld at $any$ point during execution.
